---
title: "Kajaranian dialect usage in Kajaran"
output: word_document
---
## There is a very interesting dialect situation in the Armenian language. Trying to find out the level of usage the Kajaranian dialect we have conducted a sociolinguistic interview in Kajaran to collect the data where we asked several questions about using the dialect  in this or that situation(with family, kids, acquaintances or at work etc. There have been selected 40 respondents who were subdivided into groups according to their gender and age. We have 6 age groups, from 13-88 years old and two gender groups. 

## The aim of the research is to discover the level of Kajaranian dialect usage in its area.

## Hypothesis: Kajaranian citizens mainly use its diealect.

### People were offered to value the level of usage of the dialect in several situations  by points from 1 to 10.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
options(digits = 3)
library(ggplot2)
library(dplyr)
library(reshape2)
library(knitr)
```

## Data loading, conducting data 

```{r}
data <- read.csv('C:/Users/Оник/Desktop/proj.csv', sep = ';', skip = 1, stringsAsFactors = F)
data[,10:15] <- NULL
colnames(data) <- c('name', 'gender', 'age', 'work', 'friends', 'children', 'spouse', 'parents', 'acquaintances')
data$age_group <- NA

data$age_group[data$age <= 20] <- 'teens'
data$age_group[data$age %in% 21:34] <- 'young'
data$age_group[data$age %in% 35:50] <- 'adults'
data$age_group[data$age %in% 51:65] <- 'senior'
data$age_group[data$age > 65] <- 'old'
```

## Data preprocessing, creating cumulative score variables by age and by gender

```{r}
data$children[grepl('не имеет', data$children)] <- NA
data$spouse[grepl('не имеет', data$spouse)] <- NA
data$work[grepl('не имеет', data$work)] <- NA

data$gender <- as.factor(data$gender)
data$age_group <- as.factor(data$age_group)
data$age_group <- factor(data$age_group, levels=c('teens', 'young', 'adults', 'senior', 'old'))
# typeof(data$age)
data[, 4:9] <- lapply(data[, 4:9], as.integer)
data$points <- rowSums(data[, 4:9], na.rm = T)
```

## calculating a number of points for each person in precents ( 0 min - 100%max)

```{r}
data$percent <- 100 * rowSums(data[, 4:9], na.rm = T) / (apply(data[, 4:9], 1, function(x) sum(is.na(x) == F)) * 10)
```

## Calculating a number of point for each age group 

```{r}
for (i in levels(data$age_group)){
  data$cum_age[data$age_group == i] <- sum(data$points[data$age_group == i])
  data$cum_age_max[data$age_group == i] <- sum(apply(data[data$age_group == i, 4:9], 1, function(x) sum(is.na(x) == F)) * 10)
}
data$percent_age <- data$cum_age / data$cum_age_max * 100
```

## calculating a number of points for each age-gender group

```{r}
for (i in levels(data$age_group)){
  for(j in levels(data$gender)){
    data$cum_age_gender[data$age_group == i & data$gender == j] <- sum(data$points[data$age_group == i & data$gender == j])
    data$cum_age_gender_max[data$age_group == i & data$gender == j] <- sum(apply(data[data$age_group == i & data$gender == j, 4:9], 1, function(x) sum(is.na(x) == F)) * 10)
  }
}
data$percent_age_gender <- data$cum_age_gender / data$cum_age_gender_max * 100
```

## calculating a number of points for each gender groups

```{r}
for(j in levels(data$gender)){
  data$cum_gender[data$gender == j] <- sum(data$points[data$gender == j])
  data$cum_gender_max[data$gender == j] <- sum(apply(data[data$gender == j, 4:9], 1, function(x) sum(is.na(x) == F)) * 10)
}
data$percent_gender <- data$cum_gender / data$cum_gender_max * 100
```


## creating a table of age group results 

```{r}
results_group <- subset(data, select = c(age_group, gender, cum_age, percent_age, cum_gender, percent_gender, cum_age_gender,  percent_age_gender))
results_group <- aggregate(.~ gender + age_group, results_group, function(x) x[1])

results_group
```


# Activity by age groups

```{r}
for (i in levels(data$age_group)){
  data$work_group[data$age_group == i] <- sum(data$work[data$age_group == i], na.rm = T)/sum(data$points[data$age_group == i], na.rm = T)*100
  data$friends_group[data$age_group == i] <- sum(data$friends[data$age_group == i], na.rm = T)/sum(data$points[data$age_group == i], na.rm = T)*100
  data$children_group[data$age_group == i] <- sum(data$children[data$age_group == i], na.rm = T)/sum(data$points[data$age_group == i], na.rm = T)*100
  data$spouse_group[data$age_group == i] <- sum(data$spouse[data$age_group == i], na.rm = T)/sum(data$points[data$age_group == i], na.rm = T)*100
  data$parents_group[data$age_group == i] <- sum(data$parents[data$age_group == i], na.rm = T)/sum(data$points[data$age_group == i], na.rm = T)*100
  data$acquaintances_group[data$age_group == i] <- sum(data$acquaintances[data$age_group == i], na.rm = T)/sum(data$points[data$age_group == i], na.rm = T)*100
}
```


# Activity by gender-age groups

```{r}
for (i in levels(data$age_group)){
  for(j in levels(data$gender)){
    data$work_group_gender[data$age_group == i & data$gender == j] <-
      sum(data$work[data$age_group == i & data$gender == j], na.rm = T)/sum(data$points[data$age_group == i & data$gender == j])*100
    data$friends_group_gender[data$age_group == i & data$gender == j] <-
      sum(data$friends[data$age_group == i & data$gender == j], na.rm = T)/sum(data$points[data$age_group == i & data$gender == j])*100
    data$children_group_gender[data$age_group == i & data$gender == j] <-
      sum(data$children[data$age_group == i & data$gender == j], na.rm = T)/sum(data$points[data$age_group == i & data$gender == j])*100
    data$spouse_group_gender[data$age_group == i & data$gender == j] <-
      sum(data$spouse[data$age_group == i & data$gender == j], na.rm = T)/sum(data$points[data$age_group == i & data$gender == j])*100
    data$parents_group_gender[data$age_group == i & data$gender == j] <-
      sum(data$parents[data$age_group == i & data$gender == j], na.rm = T)/sum(data$points[data$age_group == i & data$gender == j])*100
    data$acquaintances_group_gender[data$age_group == i & data$gender == j] <-
      sum(data$acquaintances[data$age_group == i & data$gender == j], na.rm = T)/sum(data$points[data$age_group == i & data$gender == j])*100
  }
}
```

# Creating a summarized table of points by activities

```{r}
results_activity <- subset(data, select = c(age_group, gender, work_group, friends_group, children_group, spouse_group, parents_group, acquaintances_group,
                                            work_group_gender, friends_group_gender, children_group_gender, spouse_group_gender, parents_group_gender, acquaintances_group_gender))
results_activity <- aggregate(.~ gender + age_group, results_activity, function(x) x[1])

results_activity
```

# ANOVA test

H0: ...

```{r}
ggplot(data, aes(age_group, percent))+
  geom_point(aes(color = age_group))

ggplot(data, aes(age_group, percent))+
  geom_boxplot(aes(fill = age_group))+
  geom_point()

kruskal_test <- kruskal.test(percent ~ age_group, data)
kruskal_test
```
## It is impossible to use ANOVA test as we see after H0 testing. But we also can make a conclusion about the tendecy we see here. The "old" group use Kajaranian dialect more than others (about 85 percent). According the median we may say - the older you are the more dialect you use.

## Scatter plot of percentage by age-gender groups

```{r}
ggplot(data, aes(age_group, percent))+
  #geom_boxplot(aes(fill = age_group))+
  geom_point(aes(color = age_group))+
  facet_grid(~gender)
```

# Student test for  scores distribution difference between men and women
# Firstly, test let's check whether the distributions are gaussian

```{r}
# Testing whether the distributions are gaussian
hist(data$percent[data$gender == 'муж.'])
shapiro.test(data$percent[data$gender == 'муж.'])
shapiro.test(data$percent[data$gender == 'жен.'])
```

Since the distributions are not gaussian,

# Wilcoxon test (non-parametric analog of t.test) - for non-gaussian distribution

```{r}

ggplot(data, aes(gender, percent))+
  geom_boxplot(aes(fill = gender))

wilcox.test(percent ~ gender, data)
```
## After H0 we see here there are enough respodents to make a welcox test. there is a difference in usage the dialect between male and female. Men use the dialect more than women in their speech.

# Activity plots


```{r}
results_activity_1 <- subset(data, select = c(age_group, gender, work_group, friends_group, children_group, spouse_group, parents_group, acquaintances_group))
results_activity_1 <- aggregate(.~ gender + age_group, results_activity_1, function(x) x[1])

long_result_activity <- melt(results_activity_1, id.vars=c('gender', 'age_group'))

ggplot(long_result_activity, aes(age_group, value))+
  geom_col(aes(fill = variable))+
  scale_fill_brewer(palette="Spectral")+
  scale_x_discrete('Age group')+
  scale_y_continuous('Percentage')+
  ggtitle('Distribution of activities by age groups')

# long_result_activity_all <- melt(results_activity, id.vars=c('gender', 'age_group'))
# 
# ggplot(long_result_activity_all, aes(age_group, value))+
#   geom_col(aes(fill = variable), position = position_fill())+
#   scale_x_discrete('Age-gender group')+
#   scale_y_continuous('Percentage')+
#   ggtitle('Distribution of activities by age-gender groups')
```
## 1. It is impossible to analyze the groups evenly cause of the absence of children and spouse. 
## 2. We also see that results of the groups are pretty the same.

## Our projects shows the result of Hypothesis: Kajaranian citizens mainly use its diealect. 
```{r}
mean(data$percent)
```
## done :)
